\section[Locally Isomorphic]{Locally Isomorphic Groups}

The theory of representations has particularly simple results for a certain class of groups, the compact groups. The property of compactness also simplifies the derivation of some general theorems.In this section we will define a compact group.

It may happen that two groups which are-essentially different appear to be identical if one is permitted to view only a part of them. We will give a rather simple example of this situation. We consider again the group R of all real numbers with addition being the operation of group multiplication. This group has a normal subgroup, the group Z of all integers. The factor group R/ Z, whose elements are cosets of Z, consists of sets of numbers such that the difference of any two of the numbers in a set is an integer. Each element can, therefore, be characterized by a number p satisfying -1/2 < p ::s 1/2. The product of two elements of R/Z corresponding to p and q correspondstop+qif-1/2<P+q ::s1/2;ifP+qisoutsidethisinterval it can be reduced by ± 1 to bring it back into the interval. The groups R and R/Z are certainly different; for example, R/Z is compact (itis isomorphic to the group of rotations in the plane) whereas R is not. The two groups are identical, however, if one considers only elements corresponding to Ipl < O.I.The groups R and R/Z are s aid to be loc ally isomorphic. This property can be defined as follows. Lie groups G and G' are locally isomorphic if there exist neighborhoods of the identity in each group and parametrizations such that the product functions are equal if their arguments are in these neighborhoods.

\endinput  % ---------------------------------------------------------------------